---
title: "US Beers and Breweries"
author: "Alvarado, Cynthia"
date: "July 1, 2017"
<<<<<<< HEAD
output:
  html_document:
    keep_md: true
=======
output: html_document:
          keep_md: true
>>>>>>> bedc972652edb55c96f1995ef33932306f0208fa
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
<<<<<<< HEAD

=======
>>>>>>> bedc972652edb55c96f1995ef33932306f0208fa
```

## Setup by Import and housekeeping

Let's import the Beers and Brewerey data stored in GitHub in CSV format.  After the import, we'll use the str() function to check that the data was loaded correctly.
<br>

The non-numeric columns were brought in as <i>factors</i>.  Using the as.character() function, they are converted to character, to make it easier to store as a string vector for graphs.


```{r gather}

#install libraries used
<<<<<<< HEAD
library(ggplot2)
library(plyr)
=======
install.packages("ggplot2")
library(ggplot2)
>>>>>>> bedc972652edb55c96f1995ef33932306f0208fa


#create variables for the url for the data we are utlilizing
brewery.url <- 'https://raw.githubusercontent.com/alvaradoc/US_Beers-Brew/master/Data/Breweries.csv'
beer.url <- 'https://raw.githubusercontent.com/alvaradoc/US_Beers-Brew/master/Data/Beers.csv'

#import the data into R
breweries <- read.csv(brewery.url, header = TRUE)
beers <- read.csv(beer.url, header = TRUE)

#check the type of data and ensure successful import
str(breweries)


#change data types from factors to characters for breweries
breweries$Name <- as.character(breweries$Name)
breweries$City <- as.character(breweries$City)
breweries$State <- as.character(breweries$State)

#check the data types for beers and ensure successful import
str(beers)

#change data types from factors to characters for beers
beers$Name <- as.character(beers$Name)
beers$Style <- as.character(beers$Style)
```

## Number of Breweries per State

To get a preliminary overview of breweries across the United States, we'll summarize the number of breweries per state. Note that there are 51 observations; DC is included on the list as a separate state.

```{r tallyByState}

#table of number of breweries per state

count(breweries, vars = "breweries$State")
```

Based on the table, Colorado has the most breweries, followed by California, Michigan, Oregon then Texas.

## Combining the datasets for further analysis

Next, we'll merge the data together to do further analysis on the data contained in the beers dataset, and summarize by state. <br>

The column in common which we will use to combine is the <i>Brewery_id</i>.  Note, however, that the column name is mis-matched. The colnames() function can be used to change the column in the Breweries data, so that it is a match. The column name "Name" is used in both datasets.  To avoid having the system rename to Name.x and Name.y, we will use the same function to rename the "Name" column in breweries.

```{r combine}

#Rename Brew_id to Brewery_Id to merge with Beer dataset
colnames(breweries)[1] <- "Brewery_id"

#Rename 'Name' in Breweries data to be unique for merge
colnames(breweries)[2] <- "Brewery"

#merge two datasets into one
brew.beer <- merge.data.frame(beers, breweries, by = "Brewery_id")
```

## Check the Merged Data

Checking the first and last six observations, to ensure that the merge is successful.  Note that there are no duplicates.  The breweries appear duplicated, but because there are breweries that produce more than one beer, the beer is unique, while the Brewerey name, city and state appear in more than one row.  That is the result we want.

```{r HT}
#view first six rows of our combined dataset
head(brew.beer)

#view last six rows of our combined dataset
tail(brew.beer)
```

## Summary of combined Data

Now that we have the data combined, and checked that it was successful, we'll run a summary, taking note of the "NA" values.  For consistency, the observations with blank ('') beer style will be changed to "NA."

```{r summry}

#we found rows to change for Style
brew.beer$Style[brew.beer$Style == ''] <- NA

#get the summary for the dataset
summary(brew.beer)
```

There are 62 NAs for ABV, 1005 NAs for IBU and finally, 5 NA values for Style, according to the summary.

## Bar chart of Median ABV and IBU by State

Below, are two bar charts for median values of Alcohol by Volume (ABV), and the International Bitterness Unit (IBU), by state.

```{r medians}
#attach dataset to refer to column names
attach(brew.beer)

#first, we calculate the Median for both
medians <- aggregate(brew.beer[, 4:5], by = list(State), FUN = median, na.rm = TRUE, na.action = 0)

#Plot Median ABV by State
ggplot(data = medians, aes(x = Group.1, y = ABV, width = 0.5)) + geom_bar(stat = "identity", position = "identity") + labs(x = "State", y = "ABV")

#Plot Median IBU by State
ggplot(data = medians, aes(x = Group.1, y = IBU, width = 0.5)) + geom_bar(stat = "identity", position = "identity") + labs(x = "State", y = "IBU")

```
<<<<<<< HEAD
<br><br>From the ABV bar graph, there are several states with high median Alcohol content: DC, Kentucky, Michigan, New Mexico and West Virginia.  The lowest median is Utah.<br>
=======
From the ABV bar graph, there are several states with high median Alcohol content: DC, Kentucky, Michigan, New Mexico and West Virginia.  The lowest median is Utah.<br>
>>>>>>> bedc972652edb55c96f1995ef33932306f0208fa

In the IBU chart, only two states show a high median value for bitterness: Maine and West Virginia. Among the states with a low bitterness value are Arizona, Kansas, Wisconsin and Wyoming.<br>

South Dakota did not have any observations for bitterness values, so it has no bar.

## Top ABV and IBU

Now we examine which states have the highest ABV percentage, and IBU value.  From the summary table ouput above, we know that the largest ABV percentage is 0.128 and the largest value for IBU is 138.

```{r max}

#Here, we determine which state has the beer with the highest ABV
State[which.max(ABV)]

#Conversely, we find the state whose beer has the highest IBU
State[which.max(IBU)]
```

Interstingly, although Colorado was not among the states with a high median for ABV, it is the state that produces the beer with highest ABV of 12.8%.  Oregon produces the most bitter beer in the US with an IBU of 138.

## Alcohol Content Summary

<<<<<<< HEAD
For full information purposes, the 6 number summary statistics for Alcohol by Volume is shown below:
=======
For full information purposes, the 5 number summary for Alcohol by Volume is shown below:
>>>>>>> bedc972652edb55c96f1995ef33932306f0208fa

```{r ABVsummy}
#We obtain the summary statistics for ABV
summary(ABV)
```

## ABV and IBU relationship

<<<<<<< HEAD
Finally, a plot of Alcohol by Volume and International Bitterness Units will show if there is a correlation.  If there is a correlation, then the plot will show how strong it is, and if it is positive or negative.
=======
Finally, a plot of Alcohol by Volume and International Bitterness Units will show if there is a correlation.  If there is a correlation, then the plot will show how strong it is and if it is positive or negative.
>>>>>>> bedc972652edb55c96f1995ef33932306f0208fa

```{r corr}
#Now we seek to find if a relationship between bitterness and alcohol content exist
ggplot(data = brew.beer, aes(ABV, IBU)) + geom_point() + geom_smooth(method = lm, se = FALSE)
```

## Conclusion

<<<<<<< HEAD
Based on the data and the scatter plot, there shows a strong positive relationship between the Alcohol content of a beer, to the bitterness unit.  In general, the higher the alcohol content, the more bitter the beer. There were 1,005 observations that were excluded.  These were observations where IBU was not available for a particular beer. <br>
The plot also shows a high density of beers with an ABV percent of 0.05, and a bitterness unit between 30 to 40.  This did not affect the correlation.  However, it does show that the majority of beers from the U.S. are mild bitter, with a 5% Alcohol by Volume. 
=======
Based on the data and the scatter plot, there shows a strong relationship between the Alcohol content of a beer, to the bitterness unit.  In general, the higher the alcohol content, the higher the bitterness of the beer. 
>>>>>>> bedc972652edb55c96f1995ef33932306f0208fa
